Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2023-02-14 analysis/paper/paper.qmd
update dockerfile

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making. The four assemblages include three from Suyanggae (SYG) and one from Yongsandong (YS). At this finer scale of resolution the picture is more complex. Phase 2 shows both highly correlated attributes, from SYG1_2, and relatively uncorrelated attributes, from SYG6_2. The two assemblages from Phase 3 show similarly correlated attributes, with the median right around our threshold value of `r threshold_value_cor` .

@@

To examine the impact of raw materials on the shape of stemmed points, we computed the correlation for different raw materials. We excluded raw materials that were used for less than three points. @fig-corr-raw-ass shows that the distribution of correlation coefficients is various depending on raw materials. Overall, the materials that used for more and equal to three points have high correlation coefficients. The stemmed points made from quartzite have highest correlation coefficients with narrow distribution. Acidic volcanic rocks are one of the highly correlated materials and individual correlations are skewed into two groups. Some of the correlations are even negatively related. Correlation values for porphyry stemmed points have a narrow distribution. Rhyolite stemmed points have wide distribution and some of the attributes are negatively related. Like acidic volcanic rocks, the correlation of shale stemmed points is also skewed into two groups. Tuff stemmed points have wide distribution with higher median value.

>>>>

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making. The four assemblages include three from Suyanggae (SYG) and one from Yongsandong (YS). At this finer scale of resolution the picture is more complex. Phase 2 shows both highly correlated attributes, from SYG1_2, and relatively uncorrelated attributes, from SYG6_2. The two assemblages from Phase 3 show similarly highly correlated attributes, with the median right around our threshold value of `r threshold_value_cor`, however the spread of correlations is high, suggesting a mixture of transmission biases.

@@

To examine the impact of raw materials on the shape of stemmed points, we computed the correlations of attributes among raw material groups. We excluded raw materials that were used for less than three points. @fig-corr-raw-ass shows that the most abundant raw materials have relatively high correlation coefficients. The stemmed points made from quartzite and porphyry have the highest correlation coefficients with narrow distributions. Acidic volcanic rocks are one of the highly correlated materials, with individual correlations divided into two groups. Similarly, the correlation of shale stemmed points is also divided into two groups. Rhyolite stemmed points have a wide distribution of correlations, and some of the attributes are negatively correlated. Tuff stemmed points have wide distribution with a high median value.
